EEG Waveform Analysis of P300 ERP with applications to Brain Computer Interfaces

### 2018-10-25 BrainScience/brainscience.tex
Small changes

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{The Electroencephalography (EEG) is not just a mere clinical tool anymore.  It has become the de-facto mobile, portable, non-invasive brain imaging sensor to harness brain information in real time.  It is now being used to translate or decode brain signals to diagnose disease or implement Human Computer Interaction devices.  The automatic processing approach which is based on using quantitative algorithms to detect the cloaked information buried in the signal, neglected the research done by the clinical EEG community.  Clinical EEG is based intensively on waveforms and the structure of signal plots.  Hence, the purpose of this work is to establish a bridge to fill this gap by reviewing and describing the procedures that have been used to detect patterns in the electroencephalographic waveforms, benchmarking them on a controlled pseudo-real dataset of an EEG Event Related Potentials waveform and verifying their performance on a public dataset of a BCI Competition.}

@@

% Keywords
\keyword{electroencephalography; EEG; waveform; ERP ;BCI;P300}

@@

EEG sensors are wearable~\citep{Puce2017} non-invasive, portable and mobile~\citep{DeVos2014}, with excellent temporal resolution, and acceptable spatial resolution~\citep{Hartman2005}.  This humble diagnosis device is been transformed into currently the best approach to detect, out-of-the lab in an ambulatory context, information from the Central Nervous System and to use that information to volitionally drive cars, steer drones, write emails, or control wheelchairs, alcohol consumption monitoring~\citep{Yuste2017,KevricJasmin2016,Tzimourta2018,Cruz2018}.

The clinical and historical tactic to analyze EEG signals were based on detecting visual patterns out of the EEG trace or polygraph\citep{Hartman2005}: multichannel signals were extracted and continuously plotted over a piece of paper. Electroencephalographers or Electroencephalography technician have decoded and detected patterns along the signals by visually inspecting them~\citep{Schomer2010}.   Nowadays clinical EEG still remains a visually interpreted test~\citep{Hartman2005}.

In contrast, automatic processing, or quantitative EEG (qEEG), was based first on analog electronic devices~\citep{Jestico1977} and later on computerized digital processing methods~\citep{Jansen1991}.  They implemented algorithmically complex procedures to decode the information associated with a clinical condition, or even some higher cognitive state \citep{Yuste2017}.  The best materialization of the automatic processing of EEG signals rests in the BCI discipline, where around $71.2\%$ is based on noninvasive EEG~\citep{Guger2017}.

%rich clinical literature

Hence, the traditional and knowledgeable approach was mainly overshadow particularly in BCI Research, and the waveform of the EEG was replaced by pragmatic procedures that were difficult to link to existing clinical EEG knowledge.  We aim to help fix this gap by providing a review of the methods which emphasize the waveform, the shape of the EEG signal and which can decode them in a supervised and semi-automated procedure.

>>>>

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{The Electroencephalography (EEG) is not just a mere clinical tool anymore.  It has become the de-facto mobile, portable, non-invasive brain imaging sensor to harness brain information in real time.  It is now being used to translate or decode brain signals to diagnose disease or implement Brain Computer Interface (BCI) devices to transmit volitional information.  The automatic decoding is mainly implemented by using quantitative algorithms to detect the cloaked information buried in the signal.  However, clinical EEG is based intensively on waveforms and the structure of signal plots. Hence, the purpose of this work is to establish a bridge to fill this gap by reviewing and describing the procedures that have been used to detect patterns in the electroencephalographic waveforms, benchmarking them on a controlled pseudo-real dataset of an EEG Event Related Potentials waveform and verifying their performance on a public dataset of a BCI Competition.}

@@

% Keywords
\keyword{electroencephalography; BCI; EEG; waveform; ERP ;P300}

@@

EEG sensors are wearable~\citep{Puce2017} non-invasive, portable and mobile~\citep{DeVos2014}, with excellent temporal resolution, and acceptable spatial resolution~\citep{Hartman2005}.  This humble diagnosis device is been transformed into currently the best approach to detect, out-of-the lab in an ambulatory context, information from the Central Nervous System and to use that information to volitionally drive cars, steer drones, write emails, control wheelchairs or to assess alcohol consumption~\citep{Yuste2017,KevricJasmin2016,Tzimourta2018,Cruz2018}.

The clinical and historical tactic to analyze EEG signals is based on detecting visual patterns out of the EEG trace or polygraph\citep{Hartman2005}: multichannel signals are extracted and continuously plotted over a piece of paper. Electroencephalographers or Electroencephalography technician decode and detect patterns along the signals by visually inspecting them~\citep{Schomer2010}.   Nowadays clinical EEG still remains a visually interpreted test~\citep{Hartman2005}.

%In contrast, automatic processing, or quantitative EEG (qEEG), was based first on analog electronic devices~\citep{Jestico1977} and later on computerized digital processing methods~\citep{Jansen1991}.  Complex procedures to decode the information associated with a clinical condition, or even some higher cognitive state \citep{Yuste2017}.  The best materialization of the automatic processing of EEG signals rests in the BCI discipline, where around $71.2\%$ is based on noninvasive EEG~\citep{Guger2017}.

The need of quantitative procedures to automate the decoding of EEG signals has been materialized in BCI where around $71.2\%$ is based on noninvasive EEG~\citep{Guger2017}.  However, methods of decoding signals based on the detection of waveforms has been scarce. Hence, this traditional and knowledgeable approach has been neglected particularly in BCI Research. We aim to help fix this gap by providing a review of the methods which emphasize the waveform, the shape of the EEG signal and which can decode them in a supervised and semi-automated procedure.
